Three $\approx$40 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia

### 2022-03-28 papers/ms.tex
update ms

%
% Numbers are from RSG-5_auto_XYZ_vl_vb_cut.csv,
% CH-2_auto_XYZ_vl_vb_cut.csv
%
We define two sets of stars in the local vicinity of our objects of
interest.  For candidate RSG-5 members, we require:
\begin{align}
  X/{\rm pc} &\in [45, 75] \nonumber \\
  Y/{\rm pc} &\in [320, 350] \nonumber \\
  Z/{\rm pc} &\in [40, 70] \nonumber \\
  v_b/{\rm km\,s^{-1}} &\in [-4, -3] \nonumber \\
  v_{l^*}/{\rm km\,s^{-1}} &\in [4, 6] \nonumber
  %45 < X/{\rm pc} < 75 \nonumber \\
  %320 < Y/{\rm pc} < 350 \nonumber \\
  %40 < Z/{\rm pc} < 70 \nonumber \\
  %-4 < v_b/{\rm km\,s^{-1}} < -3 \nonumber \\
  %4 < v_{l^*}/{\rm km\,s^{-1}} < 6 \nonumber
\end{align}
For the diffuse stars near KOI-7368 and KOI-7913, we require
\begin{align}
  X/{\rm pc} &\in [20, 70] \nonumber \\
  Y/{\rm pc} &\in [230, 270] \nonumber \\
  Z/{\rm pc} &\in [75, 105] \nonumber \\
  v_b/{\rm km\,s^{-1}} &\in [-3.5, -1.5] \nonumber \\
  v_{l^*}/{\rm km\,s^{-1}} &\in [2, 6] \nonumber
  %20 < X/{\rm pc} < 70 \nonumber \\
  %230 < Y/{\rm pc} < 270 \nonumber \\
  %75 < Z/{\rm pc} < 105 \nonumber \\
  %-3.5 < v_b/{\rm km\,s^{-1}} < -1.5 \nonumber \\
  %2 < v_{l^*}/{\rm km\,s^{-1}} < 6, \nonumber
\end{align}
and we call this latter set of stars ``CH-2''.  These cuts yield 141
candidate RSG-5 members, and 37 candidate CH-2 members.  An important
consideration, especially for CH-2, is the
contamination rate by field stars.  We assess this in the following
section.

@@

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2c.pdf}
			\includegraphics[width=0.49\textwidth]{f2d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf The stellar groups near KOI-7368, KOI-7913, and Kepler-1643
    are $\approx$40 million years old.} 
    {\it Top row}: 
    Color-absolute magnitude diagram of candidate Cep-Her members, in
    addition to candidate members of the $\delta$~Lyr~cluster
    ($\approx38$\,Myr; \citealt{bouma_kep1627_2022}) and and the Gaia
    EDR3 Catalog of Nearby Stars (gray background).  The left and
    right columns shows stars in RSG-5 and CH-2, respectively.  The
    range of colors displayed is truncated to show whether or not
    candidate members fall on the pre-main-sequence.
    {\it Bottom row}:
    TESS and ZTF-derived stellar rotation periods, with the Pleiades
    ($\approx 112$\,Myr), and Praesepe ($\approx 650$\,Myr) shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.
    Our detection efficiency for reliable rotation periods falls off
    beyond $\bpmrpo \gtrsim 2.6$.
    {\bf todo: fix zorder markers}.  
		\label{fig:age}
	}
\end{figure*}

@@

In addition, Figure~\ref{fig:age} shows that the candidate RSG-5 and
CH-2 members overlap with the $\delta$~Lyr cluster, and are therefore
roughly the same age.  To quantify this, we use the empirical method
introduced by \citet[][see their Section~6.3]{gagne_mutau_2020}.  This
idea of the method is to fit the pre-main-sequence loci of a set of
reference clusters, and to then model the locus of the target cluster
as a linear combination of these reference cluster loci.  For our
reference clusters, we used UCL, IC\,2602, and the Pleiades, from the
memberships reported by \citet{Damiani2019} and
\cite{CantatGaudin2018a} respectively.  We adopted ages of 16\,Myr for
UCL \citep{pecaut_star_2016}, 38\,Myr for IC\,2602\footnote{ Ages for
IC\,2602 vary from 40 to 46\,Myr based on lithium-depletion-boundary
(LDB) measurements \citep{dobbie_ic_2010,randich_gaiaeso_2018}, and
from 30 to 46\,Myr based on isochronal analyses
\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.},
and 112\,Myr for the Pleiades \citep{dahm_2015}.  These assumptions
and the consequent processing steps were identical to those described
in \citet{bouma_kep1627_2022}.  The mean and standard deviation of the
resulting age posterior are $46^{+9}_{-7}$\,Myr for RSG-5, and
$36^{+10}_{-8}$\,Myr for CH-2.  For comparison, the
this procedure yields an age for the $\delta$~Lyr cluster of
$38^{+6}_{-5}$\,Myr.  The slightly older isochronal age of RSG-5 is
expected given that its locus is slightly bluer and less luminous in
the upper left panel of Figure~\ref{fig:age} relative to the
$\delta$~Lyr cluster.

\subsubsection{Stellar Rotation Periods}

>>>>

%
% Numbers are from RSG-5_auto_XYZ_vl_vb_cut.csv,
% CH-2_auto_XYZ_vl_vb_cut.csv
%
We define two sets of stars in the local vicinity of our objects of
interest.  For candidate RSG-5 members, we require:
\begin{align}
  X/{\rm pc} &\in [45, 75] \nonumber \\
  Y/{\rm pc} &\in [320, 350] \nonumber \\
  Z/{\rm pc} &\in [40, 70] \nonumber \\
  v_b/{\rm km\,s^{-1}} &\in [-4, -3] \nonumber \\
  v_{l^*}/{\rm km\,s^{-1}} &\in [4, 6] \nonumber
\end{align}
For the diffuse stars near KOI-7368 and KOI-7913, we require
\begin{align}
  X/{\rm pc} &\in [20, 70] \nonumber \\
  Y/{\rm pc} &\in [230, 270] \nonumber \\
  Z/{\rm pc} &\in [75, 105] \nonumber \\
  v_b/{\rm km\,s^{-1}} &\in [-3.5, -1.5] \nonumber \\
  v_{l^*}/{\rm km\,s^{-1}} &\in [2, 6] \nonumber
\end{align}
and we call this latter set of stars ``CH-2''.  These cuts yield 141
candidate RSG-5 members, and 37 candidate CH-2 members.  An important
consideration, especially for CH-2, is the
contamination rate by field stars.  We assess this in the following
section.

@@

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2c.pdf}
			\includegraphics[width=0.49\textwidth]{f2d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf The stellar groups near KOI-7368, KOI-7913, and Kepler-1643
    are $\approx$40 million years old.} 
    {\it Top row}: 
    Color-absolute magnitude diagram of candidate Cep-Her members, in
    addition to candidate members of the $\delta$~Lyr~cluster
    ($\approx38$\,Myr; \citealt{bouma_kep1627_2022}) and and the Gaia
    EDR3 Catalog of Nearby Stars (gray background).  The left and
    right columns shows stars in RSG-5 and CH-2, respectively.  The
    range of colors is truncated to emphasize the pre-main-sequence.
    {\it Bottom row}:
    TESS and ZTF-derived stellar rotation periods, with the Pleiades
    ($\approx 112$\,Myr) and Praesepe ($\approx 650$\,Myr) shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.
    The detection efficiency for reliable rotation periods falls off
    beyond $\bpmrpo \gtrsim 2.6$.
	\label{fig:age}
	}
\end{figure*}

@@

In addition, Figure~\ref{fig:age} shows that the candidate RSG-5 and
CH-2 members overlap with the $\delta$~Lyr cluster, and are therefore
roughly the same age.  To quantify this, we use the empirical method
introduced by \citet[][see their Section~6.3]{gagne_mutau_2020}.  This
idea of the method is to fit the pre-main-sequence loci of a set of
reference clusters, and to then model the locus of the target cluster
as a linear combination of these reference cluster loci.  For our
reference clusters, we used UCL, IC\,2602, and the Pleiades, from the
memberships reported by \citet{Damiani2019} and
\cite{CantatGaudin2018a} respectively.  We adopted ages of 16\,Myr for
UCL \citep{pecaut_star_2016}, 38\,Myr for IC\,2602\footnote{ Ages for
IC\,2602 vary from 40 to 46\,Myr based on lithium-depletion-boundary
(LDB) measurements \citep{dobbie_ic_2010,randich_gaiaeso_2018}, and
from 30 to 46\,Myr based on isochronal analyses
\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.},
and 112\,Myr for the Pleiades \citep{dahm_2015}.  These assumptions
and the consequent processing steps taken to exclude field stars as
well as photometric and astrometric binaries were identical to those
described in \citet{bouma_kep1627_2022}.  The mean and standard
deviation of the resulting age posterior are $46^{+9}_{-7}$\,Myr for
RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, the this
procedure yields an age for the $\delta$~Lyr cluster of
$38^{+6}_{-5}$\,Myr.  The slightly older isochronal age of RSG-5 is
expected given that its locus is slightly bluer and less luminous in
the upper left panel of Figure~\ref{fig:age} relative to the
$\delta$~Lyr cluster.

\subsubsection{Stellar Rotation Periods}
\label{sec:rotation}

An independent way to assess the age of the candidate cluster members
is to measure their stellar rotation periods using data available from
surveys such as TESS \citep{ricker_transiting_2015} and ZTF
\citep{bellm_zwicky_2019}.  This approach leverages a storied
tradition of measuring rotation periods for open clusters of various
ages, in order to determine the rate of stellar spin-down \citep[{\it
e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
our case are especially useful, since they provide 3 to 5 lunar months
of photometry for all of our candidate CH-2 and RSG-5 members.

We selected stars suitable for gyrochronology by requiring $\bpmrpo
\geq 0.5$ and $G<16$.  The latter cut corresponds to $\bpmrpo \lesssim
2.6$, at the relevant distances.  These cuts gave 19 stars in CH-2 and
42 stars in RSG-5.  We extracted light curves from the TESS images for
these stars using the \texttt{unpopular} package
\citep{hattorio_2021_cpm}, and regressed them against systematics with
its causal pixel model.  
{\bf For the ZTF light curves, we used the default photometry?}.
We then measured candidate rotation periods
using a Lomb-Scargle periodogram, and visually inspected them
following the methods discussed in \citet{curtis_rup147_2020}.

The lower panels of Figure~\ref{fig:age} show the results.
In RSG-5,
36/42 stars have rotation consistent with ``sub-Pleiades'' (86\%).
Given the nearly perfectly clean photometric CAMD for RSG-5, this
probably gives some indication for the fraction of stars for which we
should expect
to detect rotation periods with TESS (due to e.g., pole-on stars
having less detectable starspot modulation).

Consider stars in CH-2 with 0.5 < BP-RP0, and G < 16.
19 stars meet this criterion -- all were observed by TESS.
13/19 (68\%) have rotation that is obviously consistent with a 40 Myr
age.
4 are M-dwarfs with rotation periods between 10 and 12.5 days, for which
the age interpretation is not obvious.
L.~Rebull for instance have found numerous M-dwarfs
with 10-12 day rotation periods at ages of USco ($\sim 8$\,Myr); some
might exist at ages of LCC ($\sim 16$\,Myr)
(\citealt{rebull_rotation_2020}; Rebull et al., in prep -- {\bf see
Skumanich conf slide deck -- though beware TESS stitching systematics}).
Fr the sake of argument, let's say that this provides an independent
verification that at least two thirds of our CH-2 candidates are both
photometrically and rotationally consistent with being $\approx$40
Myr.

For RSG-5, 42 stars are in the relevant color and brgithness range.
36 have rotation consistent with ``sub-Pleiades'', i.e., 86\%.
Given the nearly perfectly clean photometric CAMD for RSG-5, this
probably gives some indication for the fraction of stars for which we
should expect
to detect rotation periods with TESS (due to e.g., pole-on stars
having less detectable starspot modulation).
